Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2021-09-28 psy_peds_2.Rmd
new daily graphics added

clinical_raw_filter <- clinical_raw %>%
  filter( len_hospitalisation > min_hosp_days) %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","during_pandemic")) %>% 
  dplyr::filter(cohort  %in% c("AllAdm"))
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
```

```{r}
all_codes <- clinical_raw_filter %>% 
  left_join( obs_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
  filter( concept_type == 'DIAG-ICD10')
```

@@

new_hospitalizations_per_month3 <- new_hospitalizations_per_day %>%
  group_by( month) %>%
  summarise(count = sum(count))

@@

>>>>

clinical_raw_filter <- clinical_raw %>%
  filter( len_hospitalisation > min_hosp_days) %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","during_pandemic")) %>% 
  dplyr::filter(cohort  %in% c("AllAdm"))

```

```{r}
all_codes <- clinical_raw_filter %>% 
  left_join( obs_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
  filter( concept_type == 'DIAG-ICD10')
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
```

@@

new_hospitalizations_per_month <- new_hospitalizations_per_day %>%
  group_by( month) %>%
  summarise(count = sum(count))

@@

#### Psychiatric related conditions vs. any other condition

##### New admissions

##### Total hospitalizations
We have estimated the total number of hospitalizations in the previous steps:
- total_hospitalization_per_day
- total_hospitalization_per_week
- total_hospitalization_per_month

We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.

```{r}
total_hospitalizations_per_day_psy <- input_data %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( date ) %>%
  summarise(count_psy = n_distinct(patient_num)) %>%
  left_join( total_hospitalizations_per_day ) %>%
  mutate( period = ifelse( date < pandemic_start_date,
                                  "before_pandemic",
                                 "during_pandemic"), 
                percentage = count_psy * 100 / count)

thpd_psyPer <- total_hospitalizations_per_day_psy %>%
  ggplot(aes(x = date, y = percentage, fill = period)) +
  #geom_line()+
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  labs(y = "Percentage",
       x = "Date (by day)",
       title = "Percentage of patients with mental health related ICD codes (daily)") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

thpd_psyCount <- total_hospitalizations_per_day_psy %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Date (by day)",
       title = "Daily patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

```

We will compare the psychiatric conditions with some other disorders (e.g, digestive, respiratory, etc.)
We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.

```{r}
total_hospitalizations_per_day_disorder_group <- all_codes_desc %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' & 
        icd_description_category %in% c("Psy","Neoplams","Diseases of the respiratory system","Diseases of the digestive system" ))%>%
    group_by( date, icd_description_category ) %>%
  summarise(counts = n_distinct(patient_num))

thpd_disorderGroup <- total_hospitalizations_per_day_disorder_group %>%
  ggplot(aes(x = date, y = counts, fill = icd_description_category, color = icd_description_category)) +
  #geom_point(alpha=0.2) +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "patient number",
       x = "days",
       title = "Total number of hospitalized patients (daily)") +   
  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
